---
title: "Teacher salary bonuses"
date: '`r format(Sys.time(), "%B %e, %Y")`'
---

Teachers in North Carolina Public schools earn a bonus of \$750 if the students in their school meet a standard called "expected growth." A summary statistic called "average growth" is computed for each school; the expected growth standard is met when this summary measure exceeds zero. (This is based on [actual econometric research](https://my.vanderbilt.edu/performanceincentives/files/2012/10/200803_Vigdor_TeacherBonusesNC.pdf).)

For this example, we'll use regression discontinuity techniques to see if failing to receive a bonus motivates teachers to improve their performance in the subsequent school year.

- Running variable = `avgrow05`
- Treatment = `expgrow`
- Outcome = `avgrow06`

```{r setup, warning=FALSE, message=FALSE}
library(tidyverse)
library(haven)
library(broom)
library(rdrobust)
library(rddensity)
library(modelsummary)

schools <- read_stata("data/nc_schools.dta")
```

# Step 1: Rule based?

Yep.

# Step 2: Fuzzy or sharp?

```{r}

ggplot(schools, aes(avgrow05, expgrow, color = avgrow05 > 0)) +
  geom_point(position = position_jitter(width = 0, height = 0.25, seed = 1234)) +
  geom_vline(xintercept = 0)

```


# Step 3: Check for discontinuity in running variable around cutpoint

```{r}

schools_cheating <- schools %>% 
  filter(avgrow05 > 0 & expgrow == 1 | avgrow05 < 0 & expgrow == 0)

```


# Step 4: Check for discontinuity in outcome across running variable

```{r}
ggplot(schools_cheating, aes(avgrow05, avgrow06)) +
  geom_point() +
  geom_vline(xintercept = 0) +
  # coord_cartesian(xlim = c(-.1, .1)) +
  geom_smooth(data = filter(schools_cheating, avgrow05 < 0)) +
  geom_smooth(data = filter(schools_cheating, avgrow05 > 0))
```

```{r}
test_density <- rddensity(schools_cheating$avgrow05, c = 0)
summary(test_density)
```

# Step 5: Measure the size of the effect

```{r}

shrunken <- schools_cheating %>% 
  filter(avgrow05 > 0.1, avgrow05 < 0.1)

model <- lm(avgrow06 ~ avgrow05 + expgrow, 
            data = shrunken)
tidy(model)
```


## Parametric estimation

## Nonparametric estimation

# Step 6: Compare all the effects
