---
title: "4: Balance and Sequentiality in Bayesian Analyses"
editor_options: 
  chunk_output_type: console
---

```{r load-packages, warning=FALSE, message=FALSE}

# Load packages
library(bayesrules)
library(tidyverse)
library(patchwork)

```

## 4.9.1 Review exercises

### Exercise 4.1 (Match the prior to the description)

Five different prior models for $\pi$ are listed below. Label each with one of these descriptors: 
- somewhat favoring $\pi$ < 0.5, 
- strongly favoring $\pi$ < 0.5, 
- centering $\pi$ on 0.5, 
- somewhat favoring $\pi$ > 0.5, 
- strongly favoring $\pi$ > 0.5.

a. Beta(1.8,1.8) - centering $\pi$ on 0.5, 
b. Beta(3,2) - somewhat favoring $\pi$ > 0.5, 
c. Beta(1,10) - somewhat favoring $\pi$ < 0.5,
d. Beta(1,3) - somewhat favoring $\pi$ > 0.5, 
e. Beta(17,2) - strongly favoring $\pi$ < 0.5, 

```{r}

plot_beta(17,2) 

```

### Exercise 4.2 (Match the plot to the code) 

Which arguments to the `plot_beta_binomial()` function generated the plot below?

alpha = 3, beta = 8, y = 2, n = 4

```{r}

plot_beta_binomial(alpha = 3, beta = 8, y = 2, n = 4)

```

### Exercise 4.3 (Choice of prior: gingko tree leaf drop) 

A ginkgo tree can grow into a majestic monument to the wonders of the natural world. One of the most notable things about ginkgo trees is that they shed all of their leaves at the same time, usually after the first frost. Randi thinks that the ginkgo tree in her local arboretum will drop all of its leaves next Monday. She asks 5 of her friends what they think about the probability ($\pi$) that this will happen. Identify some reasonable Beta priors to convey each of these beliefs.

a. Ben says that it is really unlikely.
```{r}

plot_beta(1, 10)

```

b. Albert says that he is quite unsure and hates trees. He has no idea.
```{r}

plot_beta(1, 1)

```

c. Katie gives it some thought and, based on what happened last year, thinks that there is a very high chance.
```{r}

plot_beta(20, 1)

```

d. Daryl thinks that there is a decent chance, but he is somewhat unsure.
```{r}

plot_beta(5, 3)

```

e. Scott thinks it probably won’t happen, but he’s somewhat unsure.
```{r}

plot_beta(3, 5)

```


## 4.9.2 Practice: Different priors, different posteriors

For all exercises in this section, consider the following story. The local ice cream shop is open until it runs out of ice cream for the day. It’s 2 p.m. and Chad wants to pick up an ice cream cone. He asks his coworkers about the chance ($\pi$) that the shop is still open. Their Beta priors for $\pi$ are below:

coworker  |  prior
---------------------------
Kimya     |  Beta(1, 2)
Fernando	|  Beta(0.5, 1)
Ciara	    |  Beta(3, 10)
Taylor	  |  Beta(2, 0.1)

### Exercise 4.4 (Choice of prior) 

Visualize and summarize (in words) each coworker’s prior understanding of Chad’s chances to satisfy his ice cream craving.

```{r}

k <- plot_beta(1, 2)
f <- plot_beta(.5, 1)
c <- plot_beta(3, 10)
t <- plot_beta(2, .1)

k + f + c + t

```

- K is pretty sure it is *not* open.
- F is almost certain it is *not* open.
- C is confident but not sure it is *not* open.
- T is very certain that it *is* open.

### Exercise 4.5 (Simulating the posterior) 

Chad peruses the shop’s website. On 3 of the past 7 days, they were still open at 2 p.m.. Complete the following for each of Chad’s coworkers:

- simulate their posterior model;
```{r}

k1 <- plot_beta_binomial(alpha = 1, beta = 2, y = 3, n = 7)
f1 <- plot_beta_binomial(alpha = .5, beta = 1, y = 3, n = 7)
c1 <- plot_beta_binomial(alpha = 3, beta = 10, y = 3, n = 7)
t1 <- plot_beta_binomial(alpha = 2, beta = .1, y = 3, n = 7)

k1 / f1 / c1 / t1 

```

- create a histogram for the simulated posterior; and
```{r}

```

- use the simulation to approximate the posterior mean value of $\pi$.
```{r}

```


### Exercise 4.6 (Identifying the posterior) 

Complete the following for each of Chad’s coworkers:

- identify the exact posterior model of $\pi$;
- calculate the exact posterior mean of $\pi$; and
- compare these to the simulation results in the previous exercise.





