---
title: "Problem set 1"
author: "Jamie Esmond"
output:
  pdf_document: default
  html_document: default
---


```{r load-libraries-data, message=FALSE, warning=FALSE}
library(tidyverse)

cars <- read_csv("data/cars.csv")
```



# Learning R

Tell me that you worked through the primers and videos and examples at the example page for this week:

I worked through all the primers and it took a long time, but I learned a lot.


# My first plots

Insert a chunk below and use it to create a scatterplot (hint: `geom_point()`) with displacement (`displ`) on the x-axis, city MPG (`cty`) on the y-axis, and with the points colored by drive (`drv`).

```{r My_First_Plot, fig.dim=c(7,4)}
cars %>%
  ggplot() +
    geom_point(mapping = aes(displ, cty, color = drv))
```

Insert a chunk below and use it to create a histogram (hint: `geom_histogram()`) with highway MPG (`hwy`) on the x-axis. Do not include anything on the y-axis (`geom_histogram()` will do that automatically for you). Choose an appropriate bin width. If you're brave, facet by drive (`drv`).

```{r My_Second_Plot}
cars %>%
  ggplot() +
    geom_histogram(mapping = aes(hwy), binwidth = 2, boundary = 0) +
    facet_grid( ~ drv)
```

\newpage

# My first data manipulation

Insert a chunk below and use it to calculate the average city MPG (`cty`) by class of car (`class`). This won't be a plot---it'll be a table. Hint: use a combination of `group_by()` and `summarize()`.

```{r My_First_Manipulation}
cars %>%
  group_by(class) %>%
  summarise(cty_class = round(mean(cty), 1)) %>%
  arrange(desc(cty_class))
  
```
