---
title: "trash"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

### Conviction rates

-   **Measurable definition**: Appearing before the court as a non-detained defendant impacts the rate of conviction.
-   **Ideal measurement**: Collect information regarding the detention status of defendants in court and how it correlates with conviction rates.
-   **Feasible measurement**: In this case, the ideal measurement is available from the court records.
-   **Measurement of program effect**: Bail reform allows most people arrested for a misdemeanor to be released from detention and return to the court for their next court date. If the rate of conviction is lower for non-detained defendants, this provides evidence that individuals are given more equitable proceedings when they are not brought to court directly from the jail.

```{r names}

n_judges <- 15
n_prosecutors <- 25

judgeHC_names <- randomNames(n_judges, 
                           ethnicity = sample(c(1:6), n_judges, replace = TRUE, 
                                              prob = c(.005, .01, .33, .15, .5, .005)),
                           which.names = "last",
                           return.complete.data = TRUE)
prosecutorHC_names <- randomNames(n_prosecutors, 
                           ethnicity = sample(c(1:6), n_prosecutors, replace = TRUE, 
                                              prob = c(.005, .01, .33, .15, .5, .005)),
                           which.names = "last",
                           return.complete.data = TRUE)

judgeOP_names <- randomNames(n_judges, 
                           ethnicity = sample(c(1:6), n_judges, replace = TRUE, 
                                              prob = c(.005, .01, .33, .15, .5, .005)),
                           which.names = "last",
                           return.complete.data = TRUE)
prosecutorOP_names <- randomNames(n_prosecutors, 
                           ethnicity = sample(c(1:6), n_prosecutors, replace = TRUE, 
                                              prob = c(.005, .01, .33, .15, .5, .005)),
                           which.names = "last",
                           return.complete.data = TRUE)

judgesHC <- judgeHC_names %>% 
  select(last_name) %>% 
  rename(judge = last_name) %>% 
  mutate(judge_id = 1:n_judges,
         racism_j = rtruncnorm(n_judges, 0, 1, .65, .25),
         location = "Harris County")
prosecutorsHC <- prosecutorHC_names %>% 
  select(last_name) %>% 
  rename(prosecutor = last_name) %>% 
  mutate(prosecutor_id = 1:n_prosecutors,
         racism_p = rtruncnorm(n_prosecutors, 0, 1, .65, .25),
         location = "Harris County")

judgesOP <- judgeOP_names %>% 
  select(last_name) %>% 
  rename(judge = last_name) %>% 
  mutate(judge_id = 1:n_judges,
         racism_j = rtruncnorm(n_judges, 0, 1, .65, .25),
         location = "Other Place")
prosecutorsOP <- prosecutorOP_names %>% 
  select(last_name) %>% 
  rename(prosecutor = last_name) %>% 
  mutate(prosecutor_id = 1:n_prosecutors,
         racism_p = rtruncnorm(n_prosecutors, 0, 1, .65, .25),
         location = "Other Place")

judges <- judgesHC %>% 
  rbind(judgesOP)
prosecutors <- prosecutorsHC %>% 
  rbind(prosecutorsOP)

# bail <- bail %>%
#   left_join(judges, by = c("judge_id", "location")) %>%
#   left_join(prosecutors, by = c("prosecutor_id", "location")) 

```

```{r}
bail %>% 
  ggplot(aes(result, fill = trial_status)) +
    geom_bar() +
    theme_minimal() +
    theme(legend.position = "bottom")

bail %>% 
  filter(detention < 120) %>%
  ggplot(aes(ses, trial_status, color = result)) +
    geom_point(alpha = .75, size = 4,
               position = position_jitter(width = 0, height = 0.45, seed = 1234)) +
    scale_color_manual(values = colors) + 
    scale_x_continuous(breaks = seq(0,1,.1)) +
    theme_minimal() +
    theme(legend.position = "bottom")


sum2 <- bail %>%
  group_by(year, month, location) %>%
  summarise(count = n()) 

sum3 <- bail %>%
  filter(result == "Guilty Plea") %>%
  group_by(year, month, location) %>%
  summarise(count_plea = n())

sum2 <- sum2 %>% 
  left_join(sum3, by = c("year", "month", "location")) 

sum2 <- sum2 %>% 
  mutate(rate = count_plea / count) %>% 
  mutate(m_num = case_when(year == 2017 ~ 6,
                           year == 2018 ~ 18,
                           year == 2019 ~ 30,
                           year == 2020 ~ 42,
                           year == 2021 ~ 54,
                           TRUE ~ 66)) %>% 
  group_by(year, month) %>% 
  mutate(m_id = group_indices()) 

sum3 <- sum2 %>%  
  group_by(year, location) %>% 
  summarise(year_mean = mean(rate)) 

sum <- sum3 %>% 
  left_join(sum2, by = c("year", "location"))


sum %>%
  ggplot(aes(m_id, rate, color = location)) +
    geom_vline(xintercept = 36, linewidth = 1) +
    geom_point(size = 3) +
    geom_line(linewidth = 1) +
    geom_point(aes(m_num, year_mean, color = location), size = 4) +
    geom_line(aes(m_num, year_mean, color = location), linewidth = 2) +
    scale_color_manual(values = colors) +
    scale_x_continuous(expand = expansion(mult = 0.01, add = 0), breaks = seq(0,72,12)) +
    scale_y_continuous(expand = expansion(mult = 0.01, add = 0)) +
    theme_minimal() +
    theme(legend.position = "bottom",
          text = element_text(family = "serif"))
# 
# saveRDS(bail, "bail.RDS")

```
