---
title: "Exercise 12"
author: "Jamie Pantazi Esmond"
date: "today"
date-format: "long"
format: 
  html: 
    toc: true
    theme: custom.scss
    fig-cap-location: "top"
    tbl-cap-location: "top"
    title-block-banner: true
    title-block-banner-color: "#744FC6"
    fig-align: "center"
  pdf: 
    toc: true
    lof: true
    mainfont: "Cambria"
    fig-cap-location: "top"
    code-line-numbers: true
    fig-align: "center"
editor_options: 
  chunk_output_type: console
---

\newpage

# Task 1: Reflection

I see that you have copied my road trip map idea for yourself (and did so much more, of course!); now I am going to be up until 4 in the morning, trying to improve mine by following your blog posts. I just need to get caught up, so just is obviously not finished work, but it is a finished assignment. The subject is simply far to interesting to me; I must stop myself before I go too far!

\newpage

# Task 2: World map

```{r load-libraries-data, warning=FALSE, message=FALSE}
library(tidyverse)
library(sf)
library(scales)
library(tigris)        # Access geographic data from the US Census
library(tidygeocoder)  # Automated geocoding
library(osrm)          # Access OSRM through R
library(ggrepel)       # Nicer non-overlapping labels
library(glue)          # Easier string interpolation
library(patchwork)     # Combine plots nicely
library(ggspatial)     # Nicer map features like scale bars
library(jsonlite)      # Read JSON files
library(patchwork)     # Combine plots nicely
library(leaflet)       # Make interactive maps
library(lutz)          # Look up time zones for lat/lon coordinates
library(gt)            # Make fancy tables
library(rcartocolor)   # Use CARTO colors (https://carto.com/carto-colors/)
library(ggnewscale)    # Use multiple scales for the same aesthetic in ggplot

# Load and clean internet user data
internet_users <- read_csv("data/share-of-individuals-using-the-internet-1990-2015.csv") %>%
  # Rename country code column to ISO_A3 so it matches what's in the Natural Earth shapefile
  rename(users = `Individuals using the Internet (% of population) (% of population)`,
         ISO_A3 = Code)

# Load world shapefile from Natural Earth
# https://www.naturalearthdata.com/downloads/110m-cultural-vectors/
world_shapes <- read_sf("data/ne_110m_admin_0_countries/ne_110m_admin_0_countries/ne_110m_admin_0_countries.shp")
```

```{r combine-shapefile-internet-users}
# Only look at 2015
users_2015 <- internet_users %>%
  filter(Year == 2015)

users_map <- world_shapes %>%
  left_join(users_2015, by = "ISO_A3") %>%
  filter(ISO_A3 != "ATA")  # No internet in Antarctica. Sorry penguins.
```

Make a map showing the proportion of individuals in each country that had access to the internet in 2015.

```{r plot-2015-internet-uses}
#| fig.cap: "Internet Users Around the World"
# TODO: Make a map of internet users with ggplot() + geom_sf()
# Important: Remember to consider the map projection

users_map %>% 
  ggplot() +
  geom_sf(aes(fill = users)) +
  # annotation_scale(location = "bl", bar_cols = c("grey20", "white"),
  #                  unit_category = "imperial") +
  coord_sf(crs = st_crs("ESRI:54030")) +  # Robinson
  scale_fill_steps(low = "#E9E4FC", high = "#6610F2", na.value = "white", 
                   breaks = c(0,25,50,75,100), label = label_percent(scale = 1)) +
  labs(fill = "% Internet Users",
       title = "Internet Users Around the World") +
  theme_void() +
  theme(legend.position = "bottom",
        plot.title = element_text(family = "serif", face = "bold", hjust = .5))

```

\newpage

# Task 3: Personal map

```{r personal-map, warning=FALSE, message=FALSE}
#| fig-cap: "Not A Finished Map..."
# TODO: Make a map

colors <- c("#7400CC", "#CC0AA4", "#0E0ACC", "#3ACC14", 
            "#CCAC14", "#0ACCC5", "#CC5514")

together <- read_rds("data/together.rds") %>% 
  st_as_sf(coords = c("lon", "lat"), crs = st_crs("EPSG:4326"))

us_states <- states(resolution = "20m", year = 2022, cb = TRUE)

my_states <- us_states %>% 
  filter((NAME %in% c("Texas", 
                      "Louisiana", 
                      "Mississippi",
                      "Alabama",
                      "Georgia",
                      "Florida",
                      "Maryland",
                      "New Mexico")))

ggplot() + 
  geom_sf(data = my_states, fill = "#CCc1DA") +
  geom_sf(data = together, aes(color = year)) +
  coord_sf(crs = st_crs("ESRI:102003")) +  # Albers
  scale_color_manual(values = colors) +
  labs(color = NULL,
       title = "Pre-Wedding Travels",
       subtitle = "I am going to update this, but tonight, I must sleep",
       caption = str_wrap("*There is a lot of points and elements to add to this, but if I even try to start right now, nothing else in my life will ever get done, but it is a goal of mine to finish updating our map with R before we are done working together.", 125)) +
  theme_void() +
  theme(legend.position = "bottom",
        plot.title = element_text(family = "serif", face = "bold"),
        plot.subtitle = element_text(family = "serif"),
        plot.caption =  element_text(family = "serif"))

```
